### 2023-02-24 doc/manuscript/translocation.qmd
Edits to formatting and text throughout

---
format: docx
editor: visual
csl: pnas.csl
bibliography: translocation.bib
---

@@

Host evolution following disease-induced declines is one mechanism by which hosts can recover [@carlson2014; @searle2020].
Evolutionary rescue describes natural host evolution whereby rapid evolutionary changes originating from variation within a host population reverse ongoing population declines [@carlson2014].
In host-pathogen systems, evolutionary rescue is more likely to occur when hosts experience repeated selection events (e.g., through re-infections), there is a persistent pathogen reservoir or frequency-dependent infection dynamics, and populations experience increased per capita birth rates at small population sizes [@golas2021].
Evolutionary changes following disease-induced declines are known from a diversity wildlife populations [@savage2016; @epstein2016; @gignoux-wolfsohn2021; @holland2022], but linking these selective responses directly to population recoveries, as necessary for evolutionary rescue, remains rare.
Important for conservation action, selection that promotes recoveries of focal populations can also provide sources of individuals with increased resistance and/or tolerance (hereafter, "resistance") to reestablish populations extirpated following disease emergence [e.g., @mendelson2019].
To date, this potentially important species recovery strategy is rarely utilized in wildlife conservation efforts [see @joseph2018 for a notable example], and studies that evaluate its potential utility are critically needed.
<!--# Roland says: Modestly speaking, when combined with the selection results in the current paper, Joseph and Knapp (2018) does provide an example. -->

There are four conditions that will often determine the feasibility of using individuals from populations that are recovering following disease-induced declines to reestablish extirpated populations.
First, populations that survived following disease arrival have evolved increased resistance.
Second, increased resistance contributes to higher individual fitness in the presence of disease and allows positive population growth.
Third, the fitness benefits gained through selection for increased resistance manifest beyond the natal environment, allowing for the establishment of populations reintroduced into novel environments.
And fourth, reintroduced populations show long-term growth and persistence in a landscape of disease.
To date, few of these conditions are documented for any single species impacted by disease, and as a result the reintroduction of resistant hosts as a means to recover species of conservation concern remains largely an untested idea [@scheele2021].

The mountain yellow-legged (MYL) frog, comprised of the sister species *Rana muscosa* and *Rana sierrae* [@vredenburg2007], is emblematic of the unprecedented global declines of amphibians caused by the recently-emerged amphibian chytrid fungus (*Batrachochytrium dendrobatidis*; [@scheele2019]).
Once the most common amphibian in the high elevation portion of the Sierra Nevada mountains (California, USA; [@grinnell1924]), this frog has disappeared from more than 90% of its historical range during the past century [@vredenburg2007].
Due to the severity of its decline and the increasing probability of extinction, both species are now listed as "endangered" under the U.S.
Endangered Species Act.
Although this decline was initiated by the introduction of non-native trout into the once-fishless portion of the Sierra Nevada [@bradford1989; @knapp2000], the arrival of *B. dendrobatidis* (Bd) in the 1960s and its subsequent spread [@vredenburg2019] caused additional widespread population extirpations [@vredenburg2010; @rachowicz2006].
These Bd-caused declines are fundamentally different from the fish-caused declines because removal of fish populations is feasible and results in the rapid recovery of frogs [@knapp2007; @vredenburg2004].
In contrast, Bd appears to persist in habitats even in the absence of amphibian hosts [@walker2007], and therefore represents a permanent alteration of invaded ecosystems that amphibians will need to overcome to reestablish persistent populations.

Despite the catastrophic impact of Bd on MYL frogs in which most Bd-naive populations are driven to extirpation following Bd arrival [@vredenburg2010], some frog populations have persisted after epizootics [@briggs2010] and are now recovering [@knapp2016].
Frogs in these recovering populations show reduced susceptibility to Bd infection, consistent with selection for more resistant genotypes [@knapp2016].
Furthermore, exploratory efforts to reintroduce frogs collected from recovering populations to sites from which frogs were previously extirpated indicate that population establishment is possible despite ongoing Bd infection [@joseph2018].
Here, we report the results of a greatly expanded MYL frog reintroduction effort, a long-term viability modeling, and an associated analysis to evaluate genomic signatures of selection following MYL frog exposure to Bd.
Collectively, the results show that MYL frogs generally meet the four conditions for reintroduction feasibility outlined above, and provide a novel example of how the reintroduction of resistant frogs can allow the large-scale recovery of this disease-impacted amphibian.
These results may have broad implications for the many species worldwide that are threatened with extinction by emerging diseases.

@@

<!--# Add Allie's selection results here -->

#### Reintroductions allow population reestablishment

@@

Reintroductions of *R. sierrae* were conducted in Yosemite National Park (California, USA) as part of an effort to reestablish populations of this endangered species.
These reintroductions involved the collection of adult frogs from large, Bd-positive donor populations that have recovered in recent years following Bd-caused declines [@knapp2016] and translocating them to nearby recipient sites.
We refer to these reintroductions as "translocations".
During the period `r min(d1$year)`--`r max(d1$year)`, we conducted `r n_distinct(d1$site_id, d1$date)` translocations of frogs to `r n_distinct(d1$site_id)` recipient sites (`r transno_min`--`r transno_max` translocations per site; @fig-yosemap) <!--# How to restart numbering for figures in SI, add "S" prefix? -->.
Each translocated cohort included `r transfrogs_min` to `r transfrogs_max` frogs collected from one of `r n_distinct(d1$donor)` donor populations (@fig-yosemap).
Estimates of frog survival during the one year following translocation indicate that survival was highly variable between recipient sites, but relatively constant within recipient sites (for the subset of sites that received multiple translocations; @fig-translocation-survival).
As such, the average survival of frogs in the first translocation to a site was an important predictor of survival of frogs in subsequent translocations to that site (@fig-transsurvival-postdens; see **Materials and Methods - Frog translocations** for analysis details).
This suggests an important effect of site characteristics and/or (site \* frog) interactions on frog survival.
In addition, 1-year survival was higher for frogs translocated later in the study period than earlier (\@tbl-viability-earlylate; Table 2) <!--# Why isn't cross-referencing working for this table? -->, which we suggest resulted primarily from our improved ability over time to choose recipient sites with higher habitat quality for *R. sierrae*.
Finally, Bd loads changed little between immediately prior to translocation versus during the one year after translocation (@fig-bdload-beforeafter; see SI for analysis details), and were nearly always well below the level indicative of severe chytridiomycosis and associated frog mortality (5.8 ITS copies on a log~10~ scale; @vredenburg2010; @joseph2018).
<!--# Add median Bd load, proportion below 5.8 copies threshold? The text repeats in part similar results in Methods -- Translocaiton. -->

To identify site and frog-specific drivers of post-translocation frog survival, we developed a multilevel model in a Bayesian framework (see **Materials and Methods - Frog translocations** for analysis details).
Our specific interest was whether Bd load had an important effect on adult survival, as would be expected if the effect of frog resistance was insufficient to overcome the negative effect of Bd.
The best model of 1-year frog survival identified several important predictors, but Bd load at the time of translocation was not among them (bd_load; @fig-survival-postdens).
Important predictors included winter severity in the year following translocation (snow_t1), site elevation (elevation), and donor population (donor_72996; @fig-cond-effects).
Males had somewhat higher survival than females, but this effect was only marginally important (sex_male; @fig-cond-effects).
In addition to Bd load, effects of frog size (length), winter severity in the year of translocation (snow_t), day of year on which a translocation was conducted (day), and translocation order (order_first) were all unimportant (@fig-survival-postdens).
In summary, our analyses show that Bd load is not a predictor of the survival of translocated frogs, consistent with frogs in recovering populations having evolved sufficient resistance and/or tolerance to minimize the adverse effects of Bd infection.
In addition, the small changes in Bd load following translocation and relatively high survival of frogs in a substantial number of translocated cohorts suggests that resistant frogs from recovering populations can maintain the benefits of resistance in non-natal habitats (i.e., low-moderate Bd loads, minimal Bd-induced mortality).
Finally, in three locations where longer mark-recapture time series allowed us to assess the survival of adults recruited to the population (i.e., progeny of translocated individuals), naturally recruited adults had equivalent or higher survival probabilities than the originally translocated adults <!--# Show results. -->.
This result shows that frog resistance in non-natal habitats is maintained across generations of frogs.

@@

#### Long-term persistence of reestablished populations

@@

@fig-selectionresults

@fig-allelefrequencies

#### Discussion

@@

### Figures and figure legends

@@

| **Parameter**                                                             | **Value**            | **Source**                                                         |
|:--------------------------|:------------------|:-------------------------|
| $\sigma_{L_1}$, Yearly survival probability of year 1 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{L_2}$, Yearly survival probability of year 2 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{L_3}$, Yearly survival probability of year 3 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{J_1}$, Yearly survival probability of year 1 juveniles           | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $\sigma_{J_2}$, Yearly survival probability of year 2 juveniles           | 0.5                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{A_R}$, Yearly survival probability of naturally recruited adults | Varies by population | Estimated from mark-recapture studies                              |
| $\sigma_{A_T}$, Yearly survival probability of translocated adults        | Varies by population | Estimated from mark-recapture studies                              |
| $p_{L_1}$, Probability of a year 1 tadpoles remaining as a tadpoles       | 1                    | Estimated from field data, observations, natural history knowledge |
| $p_{L_2}$, Probability of a year 2 tadpoles remaining as a tadpoles       | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $p_{J_1}$, Probability of a year 1 juvenile remaining as a juvenile       | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $p_F$, Probability of a adult female reproducing in a year                | 0.5                  | Could be as high at 1, based on field observations                 |
| $F$, Number of surviving eggs produced by an adult female                 | 100                  | From observations of captive frogs                                 |
| $\omega$, Probability of juvenile successfully recruiting to an adult     | Varies yearly        | Explored different values of this parameter                        |

@@

### Figure legends

### Supporting Information

@@

![For frogs translocated to each recipient site, Bd loads for the period immediately prior to translocation versus during the 1-year period after translocation. <!--# Remove gray background and gridlines -->](figures/bdload_beforeafter.png){#fig-bdload-beforeafter width="450"}

@@

### Notes (author use only - not part of manuscript)

Science (impact factor = 63.714)

-   Research articles

    -   Present a major advance.

    -   Up to approximately 4500 words, including references, notes and captions.
        Corresponds to approximately 5 printed pages in journal (approxiumately 11 manuscript pages).

    -   Up to 6 figures or tables, and about 40 references.

-   Reports

    -   New research results of broad significance.

    -   Up to \~2500 words, including references, notes and captions.
        Corresponds to approximately 3 printed pages in journal (approximately 6 manuscript pages).

    -   Up to 4 figures or tables, and about 30 references.

Nature (impact factor = 69.8)

-   Articles

    -   Original reports whose conclusions represent a substantial advance in understanding of an important problem and have immediate, far-reaching implications.

    -   Biological sciences papers do not normally exceed 8 pages.
        This would be approximately 4300 words and 5-6 modest display items.

    -   This [article](https://www.nature.com/articles/s41586-022-04824-9) provides a potentially useful (e.g., recent) example of formatting and structure for a paper focused on ecology/evolution

PNAS (impact factor = 12.77)

-   Research reports

    -   Original research of exceptional importance.

    -   Preferred length is 6 pages, but up to 12 pages is allowed.
        6 page article is approximately 4,000 words, 50 references, and 4 medium-size figures or tables.

Nature Communications (impact factor = 17.694)

-   Article

    -   Papers published by the journal represent important advances of significance to specialists within each field.
        Open access.

    -   The main text (not including abstract, Methods, References and figure legends) is limited to 5,000 words.
        The maximum title length is 15 words.

Science Advances (impact factor = 14.957)

-   Publishes impactful research papers and reviews in any area of science.
    Journal extends the capacity of *Science* magazine to identify and promote significant advances in science and engineering.
    Open access.

-   Research article

    -   Should present a major advance.

    -   Articles can be short of long format, up to 15,000 words, with up to 10 figures and/or tables.

## Other introduction material for quick reference

P5:

we show using amphibian-pathogen system to demonstrate that resistance first demonstrate that first document evidence of strong selection in host populations that are persisting with a deadly fungal pathogen, indicative that evolutionary rescue in response to previous pathogen invasion may have occurred.
We then repeatedly translocate these "rescued" populations to non-natal habitats on the landscape and use longitudinal surveys, mark-recapture analysis, and population modeling to demonstrate that "rescued" populations recover and persist across a landscape of disease.
Thus, natural selection combined with translocations promotes the recovery of a species on the brink of disease-induced extinction.
These results may have broad implications for the many species worldwide that are threatened with extinction by the emergence of novel diseases.

Host evolution and population persistence/recovery

-   Species recovery in the face of severe disease (population growth rate \> 1 ) will depend on the ability of a species to move or change.

-   Despite several mechanisms by which recovery could occur (dispersal, evolution, phenotypic plasticity), population recovery following severe disease-induced declines has rarely (?) been documented in wildlife populations (Rabbits and myxomatosis in Australia, rabbits and hemorrhagic disease in AU (Mutze et al. 2015); trout and whirling disease: Miller and Vincent (2008); frogs in Australia, Central America, Sierra Nevada; sea stars: Moritsch and Raimondi 2018, other examples?
    Restrict to species of conservation concern?)
    .

-   Given the urgent need not to not only increase the size of remnant populations, but also reestablish extirpated populations, methods are critically needed by which the recovery of populations can be leveraged to facilitate landscape-scale recovery in the presence of ongoing disease.
    For example, the disease-caused decline of a wildlife species might ideally be followed by the eventual recovery of remnant populations via evolutionary rescue.
    These populations could then be used as sources of resistant individuals to reestablish populations extirpated following disease emergence (Mendelson et al. 2019, Joseph and Knapp 2018).

Population reestablishment using reintroduction of resistant hosts

-   "Rescued" populations provide a source of animals from which to repatriate extirpated populations and disentangle the effects of the local environment and genetics on host recoveries.

-   For rescued populations to contribute to landscape-scale species recovery, the fitness benefits gained through selection should manifest beyond the natal environment, allowing for the long-term growth and persistence of incipient populations in novel environmental conditions.

-   In this study, we first document evidence of strong selection in host populations that are persisting with a deadly fungal pathogen, indicative that evolutionary rescue in response to previous pathogen invasion may have occurred.
    We then repeatedly translocate these "rescued" populations to non-natal habitats on the landscape and use longitudinal surveys, mark-recapture analysis, and population modeling to demonstrate that "rescued" populations recover and persist across a landscape of disease.
    Thus, natural selection combined with translocations promotes the recovery of a species on the brink of disease-induced extinction.
    These results may have broad implications for the many species worldwide that are threatened with extinction by the emergence of novel diseases.

Description of MYL frog species complex, impacts of Bd, and evidence of recent recovery.

<!--# I'm not sure the following text still makes sense to include. If we de-emphasize evolutionary rescue, it seems a stretch to use it as the overarching framework for the paper. The following is my best attempt at doing so. I've left the additional text on evolutionary rescue below for use elsewhere in the manuscript. -->

We hypothesize that MYL frogs are recovering due to evolutionary rescue, that the resistant animals in these populations could be used successfully in translocations to reestablish extirpated populations, and that reestablished populations would have a relatively high probability of long-term persistence.
These hypotheses generate the following expectations:

-   Differences in allele frequencies exist between Bd-naive and recovering populations

-   Shared selective response across recovering populations

-   Changes in population growth rate due to evolution extend beyond the natal environment

-   In populations established via translocation of frogs from recovering populations, Bd load is not an important predictor of post-translocation frog survival (in contrast to important relationship between Bd load and frog survival during Bd-caused epizootics)

-   Population viability: Yearly population growth rate of translocated populations in similar environments to the natal environment should be greater than 1, indicating the capacity of the selected frogs to recover in the presence of the pathogen.

Then show how each result is consistent with those expectations.

<!--# Extra ER text for use in manuscript -->

-   "Evolutionary rescue" describes adaptation that occurs quickly enough to prevent extinction due to maladaptation to a new environment (Carlson et al. 2014), can reverse ongoing declines in some species (Golas et al. 2021; Searle et al. 2021).

-   ER has been proposed as a mechanism of recovery in wildlife populations decimated by disease (Epstein et al. 2016, Schiebelhut et al. 2018).
    This is despite the fact that when environmental change is rapid (e.g., arrival of a novel disease), the initial level of population maladaptation will be large, leaving little time for adaptive evolution to restore positive growth: Gomulkiewicz and Holt (1995), Bell and Gonzalez (2009), Lindsey et al. (2013).

    -   Tasmanian devils: Epstein et al. (2016), Wells et al., Ecology (2019)

    -   Frogs in Australia: Grogan et al. (2018), McKnight et al. (2020), Newell et al. (2013), Scheele et al., (2014)

    -   Frogs in Panama: Voyles et al. (2018)

    -   Frogs in the Sierra Nevada: Knapp et al. (2016) @knapp2016

    -   Sea Star Wasting Disease: Menge et al., Plos One (2016), Schiebelhut et al.
        PNAS (2018)

    -   White-nose syndrome: Gignoux-Wolfsohn et al. (2021) Molecular ecology.

>>>>

---
format: docx
editor: visual
bibliography: translocation.bib
csl: pnas.csl
---

@@

Host evolution following disease-induced declines is one mechanism by which hosts can recover [@carlson2014; @searle2020].
Evolutionary rescue describes natural host evolution whereby rapid evolutionary changes originating from variation within a host population reverse ongoing population declines [@carlson2014].
In host-pathogen systems, evolutionary rescue is more likely to occur when hosts experience repeated selection events (e.g., through re-infections), there is a persistent pathogen reservoir or frequency-dependent infection dynamics, and populations experience increased per capita birth rates at small population sizes [@golas2021].
Evolutionary changes following disease-induced declines are known from a diversity wildlife populations [@savage2016; @epstein2016; @gignoux-wolfsohn2021; @holland2022], but linking these selective responses directly to population recoveries, as necessary for evolutionary rescue, remains rare.
Important for conservation action, selection that promotes recovery of focal populations can also provide sources of individuals with increased resistance and/or tolerance (hereafter, "resistance") to reestablish populations extirpated following disease emergence [e.g., @mendelson2019].
To date, the use of resistant individuals for species recovery is rare [see @joseph2018 for a notable example], and studies that evaluate its potential utility are critically needed.
<!--# Other examples in addition to Joseph and Knapp (2018)? In other taxa besides amphibians? -->

There are four conditions that will often determine the feasibility of using individuals from populations that are recovering following disease-induced declines to reestablish extirpated populations.
First, populations that survived following disease arrival have evolved increased resistance.
Second, increased resistance contributes to higher individual fitness in the presence of disease and allows positive population growth.
Third, the fitness benefits gained through selection for increased resistance manifest beyond the natal environment and are maintained across generations, allowing for the establishment of populations reintroduced into novel environments.
And fourth, reintroduced populations show long-term growth and persistence in a landscape of disease.
To date, few of these conditions are documented for any single species impacted by disease, and as a result the reintroduction of resistant hosts as a means to recover species of conservation concern remains largely an untested idea [@scheele2021].

The mountain yellow-legged (MYL) frog, comprised of the sister species *Rana muscosa* and *Rana sierrae* [@vredenburg2007], is emblematic of the unprecedented global declines of amphibians caused by the recently-emerged amphibian chytrid fungus [*Batrachochytrium dendrobatidis,* @scheele2019].
Once the most common amphibian in the high elevation portion of the Sierra Nevada mountains [California, USA, @grinnell1924], this frog has disappeared from more than 90% of its historical range during the past century [@vredenburg2007].
Due to the severity of its decline and the increasing probability of extinction, both species are now listed as "endangered" under the U.S.
Endangered Species Act.
Although this decline was initiated by the introduction of non-native trout into the once-fishless portion of the Sierra Nevada [@bradford1989; @knapp2000], the arrival of *B. dendrobatidis* (Bd) in the 1960s and its subsequent spread [@vredenburg2019] caused additional widespread population extirpations [@vredenburg2010; @rachowicz2006].
These Bd-caused declines are fundamentally different from the fish-caused declines because removal of fish populations is feasible and results in the rapid recovery of frogs [@knapp2007; @vredenburg2004].
In contrast, Bd appears to persist in habitats even in the absence of amphibian hosts [@walker2007], and therefore represents a permanent alteration of invaded ecosystems that amphibians will need to overcome to reestablish populations.

Despite the catastrophic impact of Bd on MYL frogs in which most Bd-naive populations are driven to extirpation following Bd arrival [@vredenburg2010], some populations have persisted after epizootics [@briggs2010] and are now recovering [@knapp2016].
Frogs in these recovering populations show reduced susceptibility to Bd infection, consistent with selection for more resistant genotypes [@knapp2016].
Furthermore, exploratory efforts to reintroduce frogs collected from recovering populations to sites from which frogs were previously extirpated indicate that population establishment is possible despite ongoing Bd infection [@joseph2018].
Here, we report the results of (i) a greatly expanded MYL frog reintroduction effort, (ii) population viability modeling, and (iii) an analysis to evaluate genomic signatures of selection following MYL frog exposure to Bd.
Collectively, the results show that MYL frogs generally meet the four conditions for reintroduction feasibility outlined above, and provide a novel example of how the reintroduction of resistant frogs can allow the large-scale recovery of this disease-impacted amphibian.
These results may have broad implications for the many species worldwide that are threatened with extinction by emerging diseases.

@@

To identify putative genes under selection and potentially related to increased frog resistance to Bd, we used exome capture methods to compare genomes of frogs in populations with different histories of Bd exposure.
Specifically, we compared genomes of frogs sampled in four populations that have never experienced a Bd-caused epizootic ("naive") versus in five populations that experienced a Bd epizootic during the past several decades and have since recovered to varying degrees (@fig-selectionresults A: SI) [@knapp2016; @vredenburg2010].
These "recovering" populations persist in an enzootic state [@briggs2010], characterized by high Bd prevalence and Bd infection intensities ("load") in adults that are typically well below the level expected to cause frog mortality [@knapp2011, @vredenburg2010].
<!--# Modify text/remove text here/elsewhere to avoid repetition of this. --> Bd-exposure histories of these populations are based on a decade or more of data from VES and Bd surveillance using skin swabbing [e.g., @knapp2016].
The comparison of frog genomes between naive and recovering populations is not as powerful as comparing samples collected from populations before and after Bd exposure, but because Bd arrived in most MYL frog populations decades ago and population persistence/recovery is rare and unpredictable, no such samples are available and are unlikely to be available in the future.
As such, the current study design is the best possible under the circumstances.
To detect regions of the genome under selection, we used a multivariate linear mixed model to compare genomes of frogs from naive and recovering populations, and created "liberal" and "stringent" sets of outlier variants using a Bonferroni-corrected significance level of 0.05 and 0.01, respectively.

The samples clustered into three separate groups in principal component space, each containing at least one population from both the naïve and persistent groups (@fig-selectionresults B: SI).
The liberal set of variants contained 38 outliers (35 SNPs and 3 structural variants) from 30 distinct genes across 16 contigs (@fig-selectionresults C, D: SI).
One gene in this group does not have an associated annotation.
No biological functions, molecular functions, or cellular processes were significantly overrepresented in the set of 29 annotated outlier genes.
The stringent set contained 11 outlier SNPs from 7 distinct genes across 4 contigs (@fig-selectionresults C, D: SI).
For these SNPs, frequency differences between the persistent and recovering groups ranged from 41% to 86%.
Most SNPs in the stringent set showed frequency differences in only a subset of the sampled populations (@fig-allelefrequencies A, B) and the patterns may be confounded by frog relatedness and/or population structure despite the methods we used to account for these factors.
However, the outlier SNP in the RIN3 gene showed consistent differences in frequencies across all populations (@fig-allelefrequencies C), suggestive of strong selection at this locus associated with exposure of frogs to Bd.
<!--# Allie should add a brief description of the function of RIN3 here. Also, emphasize key messages related to selection, resistance. And, of course, rework any of the text as necessary. -->

<!--# Figure edits:  selectionresults: Correctly locate population 7, capitalize panel labels, label liberal and stringent outlier sets. outlier map: Rearrange panels to put RIN3 last? Relocate pie charts to avoid the perception that all recovering populations are west of naive populations, add A, B, C labels to panels. -->

@@

Reintroductions of *R. sierrae* were conducted in Yosemite National Park (California, USA) as part of an effort to reestablish populations of this endangered species.
These reintroductions involved the collection of adult frogs from large, Bd-positive donor populations that have recovered in recent years following Bd-caused declines [@knapp2016] and translocating them to nearby recipient sites [@joseph2018].
We refer to these reintroductions as "translocations".
During the period `r min(d1$year)`--`r max(d1$year)`, we conducted `r n_distinct(d1$site_id, d1$date)` translocations of frogs to `r n_distinct(d1$site_id)` recipient sites (`r transno_min`--`r transno_max` translocations per site; @fig-yosemap SI) <!--# How to restart numbering for figures in SI, add "S" prefix? -->.
Each translocated cohort included `r transfrogs_min` to `r transfrogs_max` frogs collected from one of `r n_distinct(d1$donor)` donor populations (@fig-yosemap SI).
Estimates of frog survival during the one year following translocation indicate that survival was highly variable between recipient sites, but relatively constant within recipient sites (for the subset of sites that received multiple translocations; @fig-translocation-survival).
As such, the average survival of frogs in the first translocation to a site was an important predictor of survival of frogs in subsequent translocations to that site (@fig-transsurvival-postdens SI).
This suggests an important effect of site characteristics and/or (site \* frog) interactions on frog survival.
In addition, 1-year survival was higher for frogs translocated later in the study period than earlier (@tbl-viability-earlylate SI) <!--# Should be Table 1, not Table 2. Rearrange table order in manuscript. -->, which we suggest resulted primarily from our improved ability over time to choose recipient sites with higher habitat quality for *R. sierrae*.
Finally, Bd loads changed little between immediately prior to translocation versus during the one year after translocation (@fig-bdload-beforeafter SI), and were nearly always well below the level indicative of severe chytridiomycosis and associated frog mortality (5.8 ITS copies on a log~10~ scale) [@vredenburg2010; @joseph2018].
<!--# Add median Bd load, proportion below 5.8 copies threshold? The text repeats in part similar results in Methods -- Translocation. -->

To identify site and frog-specific drivers of post-translocation frog survival, we developed a multilevel model in a Bayesian framework.
Our specific interest was whether Bd load had an important effect on adult survival, as would be expected if the effect of frog resistance was insufficient to overcome the negative effect of Bd.
The best model of 1-year frog survival identified several important predictors, but Bd load at the time of translocation was not among them (bd_load; @fig-survival-postdens SI).
Important predictors included winter severity in the year following translocation (snow_t1), site elevation (elevation), and donor population (donor_72996; @fig-cond-effects).
Males had somewhat higher survival than females, but this effect was only marginally important (sex_male; @fig-cond-effects).
In addition to Bd load, effects of frog size (length), winter severity in the year of translocation (snow_t), day of year on which a translocation was conducted (day), and translocation order (order_first) were all unimportant (@fig-survival-postdens SI).
Collectively, these results show that Bd load is not a predictor of the survival of translocated frogs, consistent with frogs in recovering populations having evolved sufficient resistance to minimize the adverse effects of Bd infection.
In addition, the low-moderate Bd loads that characterized frogs following translocation and relatively high survival of frogs in the majority of more-recently translocated cohorts suggests that resistant frogs from recovering populations can maintain the benefits of resistance in non-natal habitats.
Finally, in three locations where longer mark-recapture time series allowed us to assess the survival of adults recruited to the population (i.e., progeny of translocated individuals), naturally recruited adults had equivalent or higher survival probabilities than the originally translocated adults <!--# Show results before "Collectively" above. Summarize result here, e.g., with the next sentence. -->.
Therefore, frog resistance in non-natal habitats is maintained across generations.

@@

@@

### Discussion

@@

@@

| **Parameter**                                                             | **Value**            | **Source**                                                         |
|:---------------------------|:-----------------|:-------------------------|
| $\sigma_{L_1}$, Yearly survival probability of year 1 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{L_2}$, Yearly survival probability of year 2 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{L_3}$, Yearly survival probability of year 3 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{J_1}$, Yearly survival probability of year 1 juveniles           | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $\sigma_{J_2}$, Yearly survival probability of year 2 juveniles           | 0.5                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{A_R}$, Yearly survival probability of naturally recruited adults | Varies by population | Estimated from mark-recapture studies                              |
| $\sigma_{A_T}$, Yearly survival probability of translocated adults        | Varies by population | Estimated from mark-recapture studies                              |
| $p_{L_1}$, Probability of a year 1 tadpoles remaining as a tadpoles       | 1                    | Estimated from field data, observations, natural history knowledge |
| $p_{L_2}$, Probability of a year 2 tadpoles remaining as a tadpoles       | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $p_{J_1}$, Probability of a year 1 juvenile remaining as a juvenile       | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $p_F$, Probability of a adult female reproducing in a year                | 0.5                  | Could be as high at 1, based on field observations                 |
| $F$, Number of surviving eggs produced by an adult female                 | 100                  | From observations of captive frogs                                 |
| $\omega$, Probability of juvenile successfully recruiting to an adult     | Varies yearly        | Explored different values of this parameter                        |

@@

### Figures

## Supporting Information

### Figures

@@

![For frogs translocated to each recipient site, Bd loads for the period immediately prior to translocation versus during the 1-year period after translocation.](figures/bdload_beforeafter.png){#fig-bdload-beforeafter width="450"}

@@

###
